---
title: "A Jrich worked example"
author: "D Miranda E"
date: "29/02/2016"
output: html_document
---

These functions calculate the taxonomic measures presented in Posadas et al 2001 and in Miranda-Esquivel 2016 to introduce Jack-knife re-sampling in evolutionary distinctiveness prioritization analysis. This analysis is a way to evaluate the support of the ranking of the areas to modifications in the data used, and therefore quantifying the persistence of a given area in a conservation analysis.

For further information, you could read http://link.springer.com/chapter/10.1007/978-3-319-22461-9_11



> ###  **1.** An example with a single topology and distribution to reproduce Figure 1 in Miranda-Esquivel (2016)  

First of all, we remove everything from the R environment.

```{r}
#
rm(list=ls())

```

And close the graphic devices, if there is any.

```{r}
#
if (dev.cur()!=1){dev.off()}

```

Now, we get the latest version of the library, from GitHub.

```{r,eval=FALSE}

#library("devtools")

#install_github("Dmirandae/jrich")

```

We load the library before proceeding.

```{r}

library(jrich)

```

ans set working directory to the R Data (make sure you have changed it to your own directory).

```{r}

# setwd("./myData/")


```

Now, we can read a tree from `figure1.tre`, which is written in Newick format or we can use directly the tree from the data file.


To read a file form a file use:

```{r, eval=FALSE}


tree.figure1 <- read.tree ("yourTree.tre")

```

Or, upload the data set, using the `data` funtion.

```{r}

 data(tree)

```

We can plot the tree.

```{r}

plot(tree, main= "Figure 1. Area Cladogram")

```

We can also read distributions using the species per area format. The distributions could be a csv file with several features: Each line has a species name and an area, and multiple areas for the same species means a widespread taxon in multiple lines.

The following function creates a data frame for the species distributions.

```{r, eval=FALSE}

distrib.figure1 <- Read.Data("figure1.csv")

head(distrib.figure1)

```


Also, we could read the example distribution.

```{r}

data(distribution)

```



We run the initial Index calculation, with a verbose output.

```{r}
library(jrich)

initial.Values <- Calculate.Index(tree=tree, distrib = distribution, verbose=T)

initial.Values 

```

Note that the figures for __Is/Ws__ indices here are different from Figure 1 in Miranda-Esquivel 2016, as here are re-scaled to sum 1, but the proportions are exactly the same.

To obtain the same figures for __Is/Ws__ indices as in Figure 1, we must use this code: 

```{r}

figure1.Values <-  Calculate.Index(tree=tree, distrib = distribution, verbose=F, standard = "tree")

figure1.Values

all.equal(initial.Values,figure1.Values)

```

**1.** Correlations between values

Plot the initial Values for the index that "explains" the most with this line:

```{r}

correlations <- cor(initial.Values[,2:10],initial.Values[,2:10])

```

Do not forget using the following code to avoid the "autocorrelation".

```{r}

diag(correlations) <- 0.0

```

Now we can determine the 'Best' descriptor:

```{r}

best.Index <- which.max(apply(correlations,2,sum))

best.Index

```

With the following instruction we can get the name of index that displays the highest value,
and the row of the column __rich__ in which the object is found. 

Keep in mind that richness in not a good predictor for all indices. 

```{r}

which(abs(correlations[,"rich"])==max(abs(correlations[,"rich"])))

```
 
To plot aesthetic graphics, load the library `ggplot2`.
 
```{r}

library(ggplot2)

```

**2.** The index that "explains the most", without resampling is:

```{r}

best.Index <-   which.max(apply(correlations,2,sum))

best.Index

qplot(initial.Values$area,initial.Values[,names(best.Index)], xlab = "Areas", 
      ylab =paste(names(best.Index)," values"), main = paste("Figure 2. Values of the most informative index: ",names(best.Index)," Index"))

```

In this example, Areas __A / F / G / H__ have the same ranking, as area __A__ harbors species __I__, while areas __F / G / H__ have the highest absolute richness.

We run the analysis with a single Jackknife replicate with a jtip value of 0.5. 

```{r}

jack.Values <-  Calculate.Index(tree=tree, distrib = distribution,jtip = 0.5)

```

The absolute difference between these two outputs can be computed with this instruction:

```{r}

all.equal(initial.Values, jack.Values)

```

**3.** But a single replicate is not interesting, therefore we can repeat the process 100 times, using two approaches:

  + **3.1 **Jackknifing with a jtip value of 0.5, 100 replicates, using the `Calculate.Index` function.
    
```{r}

jack.Ranking.100 <- list()
for (i in 1:100){
  ## if you want to get the output number replicate, uncomment this line:
  ##print(paste("replicate #",i), )
  
  jack.Ranking.100[[i]] <-  as.data.frame(Rank.Indices(Calculate.Index(tree=tree, distrib = distribution, verbose=FALSE,jtip = 0.5)))
}

```

+ **3.2 **Comparing the whole ranking for the best index, 100 times.

```{r}

initial.Ranking <-  as.data.frame(Rank.Indices(initial.Values))

jack.Ranking.100.comparison <- NULL

for (i in 1:100){
  
  if(!all(jack.Ranking.100[[i]][,best.Index] == "X0X")){
    jack.Ranking.100.comparison[i] <- all.equal(initial.Ranking[,best.Index], 
                                                  jack.Ranking.100[[i]][,best.Index])  
  }else jack.Ranking.100.comparison[i] <- 0
}

```

Get the number of hits or the extent of the Jackknife replicate to recover the initial ranking:

```{r}

length(which(jack.Ranking.100.comparison==TRUE))

```

To estimate the error, we use this code:

```{r}

length(which(as.data.frame(jack.Ranking.100.comparison)!=TRUE))

```

These two figures indicate that the best index is not __I__. As it is also important to estimate the error, we can use the following code to compute it.

```{r}

jack.Mismatch <- jack.Ranking.100.comparison[jack.Ranking.100.comparison!=TRUE]

count.Jack.Mismatch <- gsub(" string mismatches","",jack.Mismatch)

count.Jack.Mismatch <- as.numeric(count.Jack.Mismatch) 

```

Plot the distribution of the error, not so bell-shaped.

```{r}

hist(sort(count.Jack.Mismatch,na.last = NA), main = "Figure 3. Histogram of Error Distribution", xlab= "Jack Mistmatch count")

```

**4.** A wrap to theprevious function, and evaluating the number of times we recovered 1/2/3 position in the ranking.

Note that `Calculate.Index` recovers the index values while `Best.Index` recovers the ranking comparison.

```{r}

jack.figure1.jtip05.100replicates <- Best.Index(tree=tree, distrib = distribution, jtip = 0.5, replicates = 50, success = c(1:2))

jack.figure1.jtip05.100replicates

best.Index = names(jack.figure1.jtip05.100replicates)[c(which(jack.figure1.jtip05.100replicates == max(jack.figure1.jtip05.100replicates)))]
```

__W / Ws__ explains better than __I__, as they have a Jackknife value. In this context we can plot __Ws__.

```{r}
for (i in 1:length(best.Index)){

print(best.Index[i])

print(qplot(initial.Values$area,initial.Values[,best.Index[i]], xlab = "Areas", 
      ylab =paste(best.Index[i]," values"), main = paste("Figure 4. ",best.Index[i]," Index")))
}
```

As it is shown in __Figure 4.__, areas __F/G/H__ have a higher value as they have higher richness, but even so, the support is relatively low.

> ###  **2.** An example with two topologies and distributions  

For this example, We will work with a tree and a distribution for an real taxon: _Puranius_ (These two files could be found in in the data directory). 

```{r, eval=FALSE}

tree.Puranius <- read.tree ("puranius.tre")

distrib.Puranius <- Read.Data ("puranius.csv.gz")

```

Now we will assign a list to the object `data.Puranius` in order to join the tree and the distribution, which could certainly be made with a function as well. However, I prefer this approach. 

```{r, eval=FALSE}

data.Puranius  <- list()

data.Puranius[[1]] <- tree.Puranius 

data.Puranius[[2]] <- distrib.Puranius

head(data.Puranius)

```

We will proceed just as we did for the first Taxon, thence we got a tree and a distribution for another Taxon: _Janus_.

```{r, eval=FALSE}

tree.Janus <- read.tree ("Janus.tre")

distrib.Janus <- Read.Data("Janus.csv.gz")

data.Janus  <- list()

data.Janus[[1]] <- tree.Janus

data.Janus[[2]] <- distrib.Janus

data.Janus
```

The following code creates a list to handle multiple datasets.
In this case, the list `Multitaxon1` contains two objects.

```{r, eval=FALSE}
Multitaxon1 <- list()

Multitaxon1[[1]] <- data.Janus

Multitaxon1[[2]] <- data.Puranius

```

The function `Multi.Index.Calc` calculates indices values for a MultiData list. For this case, it computes the Initial ranking for `Multitaxon1` with default values. This action is performed just once.

```{r}

data(Multitaxon1)

initial.Values.Multi <-  Multi.Index.Calc(Multitaxon1)

```

Before plotting the initial Values for the Index that explains the most, make sure that the library `ggplot2` is loaded.

```{r}
library(ggplot2)
```

**1.** Correlations between values

Plotting as example, the index value that explains the most is:

```{r}
correlations.Multi <-  cor(initial.Values.Multi[,2:10], initial.Values.Multi[,2:10])
```

To avoid the highest "autocorrelation", use this line:

```{r}
diag(correlations.Multi) <- 0.0
```

In case we want to know the index that explains the most, use this code:

```{r}
best.Index.Multi <-   which.max(apply(correlations.Multi,2,sum))

qplot(initial.Values.Multi$area,initial.Values.Multi[,names(best.Index.Multi)], xlab = "Areas", ylab =names(best.Index.Multi), main = "Figure 5. Values of the Index that explains the most per area")
```

> **2.** A  delete experiment: Prob jtip jtopol 0.5 

```{r}
jack.Multi <-  Multi.Index.Calc(Multitaxon1,jtip = 0.5,jtopol = 0.5) 
```

See that the `jtopol` and `jtip` present the deleted proportions, and the two data frames are different.

```{r}
all.equal(jack.Multi,initial.Values.Multi)

vector <-   which.max(apply(correlations,2,sum))
```

**3.** Jack Ranking Multitaxon 100 times

This operation computes the index values and returns a data matrix. Diverse from the first method, it allows to perform this operation 100 times. The number of replicates can be easily modified, by changing the rank in for `(i in 1:100)`.

```{r}

jack.Ranking.100 <- list()

for (i in 1:100){
  print(paste("replicate #",i))
  jack.Ranking.100[[i]] <-  as.data.frame(Rank.Indices(Multi.Index.Calc(Multitaxon1,
  jtip = 0.5,jtopol = 0.5)))
}

```

We can get the ranking in the first run. It is only necessary to run this line:

```{r}

jack.Ranking.100[[1]][,vector]

```

**4.** Convert the initial values to a ranking

The initial ranking for Multitaxon is assigned as follows:

```{r}
initial.Ranking <-  as.data.frame(Rank.Indices(Multi.Index.Calc(Multitaxon1)))


initial.Ranking[,vector]
```

**5.** Compare the whole ranking for the best index, 100 times.

```{r}

jack.Ranking.100.comparison <- NULL

for (i in 1:100){
  jack.Ranking.100.comparison[i] <- all.equal(initial.Ranking[,vector], 
                                              jack.Ranking.100[[i]][,vector])
}

```

Compute the number of hits or the Jackknife replicate ability to recover the initial ranking:

```{r}
length(which(jack.Ranking.100.comparison==TRUE))
```

As in the first example, get an estimation of the error.

```{r}

length(which(as.data.frame(jack.Ranking.100.comparison)!=TRUE))

```

And with this instruction we can identify the type of error.

```{r}

jack.Mismatch <- jack.Ranking.100.comparison[jack.Ranking.100.comparison!=TRUE]

count.Jack.Mismatch <- gsub(" string mismatches","",jack.Mismatch)

count.Jack.Mismatch <- as.numeric(count.Jack.Mismatch) 

```

This code computes the distribution of the error, which will rather be bell shaped.

```{r}

hist(sort(count.Jack.Mismatch), main="Figure 6. Histogram of Error Distribution",  xlab= "Jack Mistmatch count")

```
